Denotational Semantics of General Payment Primitives, and Its Payment System

### 2022-10-03 packages/spec-haskell/yellowpaper/Paper1.tex
Real Time Balance section

%% Include a preamble at choice during processing.
\input{preamble.tex}
\begin{document}

@@

In Chapter 4, a reference implementation of the general payment primitives and its payment system
called
\textit{Superfluid Money} is introduced.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Superfluid Money - A Reference Implementation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

>>>>

%% Include the preamble at choice during processing.
\input{preamble.tex}

\begin{document}

@@

In Chapter 4, a reference implementation of the general payment primitives and its payment system
called \textit{Superfluid Protocol} is introduced.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Superfluid Protocol - A Reference Implementation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Superfluid protocol (``the protocol'')\footnote{The code base is available
at \url{https://github.com/superfluid-finance/protocol-monorepo/}.} is the first implementation of
the \textit{denotational semantics of payment primitives} (though before it was formalized and named
so) on \textit{Ethereum Virtual Machine} \cite{wood2014ethereum}. The first version of the protocol
is written in Solidity programming language \footnote{\url{https://soliditylang.org/}}.

To better serve as a reference implementation of the \textit{modern payment system} formalized by
this paper and its sequels, an implementation in Haskell was created. It aims to implement the full
specifications of (a) denotational semantics of payment primitives (b) compositional financial
contracts for the payment primitives (c) token \& note model of money mediums, and their execution
environment.

This paper covers the implementation of (a).

@@

For the purpose of compositional financial contracts, it is useful to separate monetary values that
bearers can use immediately from ones that are set aside for other financial purpose. The concept
of \textit{real time balance} is thus created. \textit{Real time balance} is a
functorful\footnote{Bartosz Milewski has an excellent series on these
topics: \url{https://bartoszmilewski.com/2015/01/20/functors/}, and he uses the term ``functorful''
to convey the idea of generalized container.} of monetary values. It can be converted to a single
monetary value.

Although since this is not the subject of this paper, it suffices to just show its definition
instead:

\begin{code}
-- Note that we omit the definition of AnyTypedValue here,
-- since it is not relevant to the idea here.

class ( MonetaryValue v
      , Foldable rtbF
      , Monoid (rtbF v)
      , Eq (rtbF v)
      ) => RealTimeBalance rtbF v | rtbF -> v where

    -- | Convert a single monetary value to a RTB value.
    valueToRTB :: Proxy rtbF -> v -> rtbF v

    -- | Net monetary value of a RTB value.
    netValueOfRTB :: rtbF v -> v
    netValueOfRTB = foldr (+) def

    -- | Convert typed values to a RTB value.
    typedValuesToRTB :: [AnyTypedValue v] -> rtbF v

    -- | Get typed values from a RTB value.
    typedValuesFromRTB :: rtbF v -> [AnyTypedValue v]
\end{code}

In the paper, we refer to the real time balance values as $rtb$, and its function $netValueOfRTB$ is
what matters the most here, it converts any $rtb$ to the \textit{monetary value} which is what the
model of money distribution needs.
